---
title: "Learning.R"
output: html_notebook
---


Vectors are data elements containing same data type.
it is a 1-d object
```{r}
x <- 3:15
x
```
x is a vector containing 10 vector elements

In R, you work with vector. It is meant to carry vectorized operations
```{r}
y <- 2
y
```
y is a vector containing single vector element

Vector type:
1. Numeric vector: int or double
2. character vector
3. logical vector

vectors are enumerated means they are listed in a specific order,
that means it allows you to call specific vector element from a vector
```{r}
x[4]
```
I am calling 4th element from a vector x.
This also tells you that vector indexation starts from 1 and not 0(unlike arrays).
int[1:13], tells you that vector is of length 13. 
meaning it contains 13 vector elements

Concept of implicit coercion:
R may force change a data type without telling you.
e.g. 
```{r}
a <- c("Hello","I","am","sleeping",3)
a
```
Since vector needs to be of a same data type. here x is a character vector, except for 3.
#so R will do implicit coercion and by that it means R will change 3, a numeric vector, into a character.
And R will do this without telling you-"implicitly".
Also
```{r}
b <- c(1:4,"hello")
b
```
R will coerice an almost numeric vector into a character vector, if you sees even a single character vector in it.
#to create vector, you use c() function- combine function

You can also check the data type of your vector
```{r}
is.numeric(x)
is.integer(x)
is.double(x)
is.character(a)
is.logical(b)
```
logical vector will be most common in use when you are working with filtering data in a matrix or data frame. 

#Three important functions
1.seq() function
?seq()
seq() is a sequence generator
```{r}
seq(from=30,to=90,length.out=5)
```
length.out is how many vector elements you want to see in your output.
```{r}
seq(from=30,to=90,by=3)
```
by parameter is the gap between vector elements
by = (from-to)/(length.out-1)
= 63/3
using seq() to create vectors
```{r}
c <- seq(50,90,length.out=10)
c
```
2. colon function
```{r}
d <- 5:17
d
```
difference between colon and seq()?
seq() provides you additional flexibility, you can decide the length of your output, you can decide jump between vector elements.
```{r}

```

3. rep() function
```{r}
?rep
rep(x=3,times=5)
rep(seq(2,10,by=1.5),length.out=15)
```
Square bracket 
This is used to access individual elements of a vector
```{r}
x[10]
x[-3]
x[1:3]
x[c(3,4,10)]
x
x[-1:-5]
z <- x[-1:-5]
z
z[9]
```
x[-3] means, all elements except the third element of that vector
x[1:3], gives you first three vectors
x[c(3,4,10)], here vector c(3,4,10), is creating an indexation. you are calling 3rd, 4th and 10th vector element
x[-1:-5], will delete first five vector elements from the vector list
z <- x[-1:-5], you can use square-bracketing to create a new vector as well.
z[9], asking a vector for an element that doesn't exist, will return you NA (meaning not available).

Vectorized operation:
Think of them as mathematical operation (adding, subracting, division, multiplication)
vectorized operation between vectors is possible if both vectors are of same length
Best part of R is mathematical operation with large vectors
```{r}
x+d
x+d
x/d
x-d
x*d
x+z
e <- 1:10
f <- 1:5
e+f
```
x+z, x and z are both of different length. R will still add vectors (by recycling of vectors, but it will show a warning message of sort,?, for recycling, larger vector must be a multiple of shorter vector)

e+f, again e and f are of different length. But R hasn't shown any warning message as e is a multiptle of f.

Vectorized vs devectorized operation:
let's take mathematical multiplication
1. Vectorized operation
```{r}
rnorm(5)
1:5
1:5 * rnorm(5)
```
2. de-vectorized operation
```{r}
g <- seq(3,15,length.out=10)
h <- rep(5,10)
k<-0
for(j in 1:10){
  k[j] <- print(g[j] * h[j])
}
k
```

so, de-vectorized operations are complex and lengthy compared to vectorized operation. And this is the reason R is perfect for statistical computing as it makes work easy.

Law of large numbers
When n approaches infinity (means your sample size is quite large), mean of the sample is equal to its fair probability.

And for normal distribution :
most values (68%) fall between (mean-sd) and (mean+sd).
```{r}
rnorm(1000)
```
rnorm() is used to get random values from a normal distribution. 
For normal distribution, mean=0 and sd=1
For 1000, large sample size, most values will fall between -1 and 1 ((0-1)and(0+1)).
```{r}
N <- 1000
counter <- 0
for(i in rnorm(N)){
  if(i < 1 && i > -1){
    counter <- counter + 1
 }
}
```

counter= 684/1000 = .684 = 68.4%
This tells us that 68.4% of the rnorm(1000) values fall between -1 and 1.

What is rnorm() ?
let's open its documentation
?rnorm
rnorm () gives you random deviates around the mean. This function comes under normal distribution.
rnorm(n), n stands for number of observation or sample size.

```{r}
rnorm(n=10,mean=13,sd=3)
```
most values of rnorm() will be between (mean-sd) and (mean+sd).

Matrices
matrices are 2-dimensional objects. 
Each point has (x,y).
Matrices and data frames are used to form tables.
Difference between matrices and data frame is that:
matrices contains elements of same type, whereas, data frame may contain mix data types.

Three functions used to create a matrix:
1. matrix() - takes a single vector and bends it into 2D object
?matrix
```{r}
x <- 1:20
matrix(x,nrow=4,ncol=5,byrow=F)
```
parameter byrow=TRUE, allows you to arrange the matrix row-wise
```{r}
A<-matrix(x,nrow=4,ncol=5,byrow=T)
```
Note: nrow x ncol = number of observations
n=20 , 5x4=20, 10x2=20

You will rarely use matrix().

2. rbind()
This allows you to combine objects row-wise.
All vectors are of same length
It is most common way to form a matrix.
Let's see how to use it:
```{r}
a <- rep(5,times=10)
b <- seq(from=200,to=300,length.out=10)
c <- 1:10
d <- c(2,6,7,8,4,2,3,4,2,1)
B <- rbind(a,b,c,d)
rownames(B) <- c("John","keth","joy","kelly")
B <- rbind(John=a,keth=b,joy=c,kelly=d)
colnames(B) <- letters[1:10]
```
John=a means you are providing rownames
Note: rownames are character types, so use double quotations.

3. cbind()- binds the vectors column-wise
All vectors are of same length
```{r}
cbind(a,b,c,d)
C <-cbind(John=a,keth=b,joy=c,kelly=d)
```
rbind() binds row-wise, cbind() binds column-wise.

Transpose of a vector
Transpose function is used to convert rows into columns.
```{r}
B
t(B)
C
t(C)
```
When to use transpose function:
matplot() is used to plot data column-wise. 
If you have your vectors arranged by rbind(), then you must use t() first, and then matplot().

How to call a vector element from a matrix
nameofthematrix[row,column]  or
nameofthematrix[rowname,columnname]
```{r}
B
B[2,3]
B["keth",3]
```
Note: you must use double-quotations when using names

Deleting rownames and columnnames
step1: check names
```{r}
rownames(B)
colnames(B)
```
step2: delete names
syntax:
rownames(nameofmatrix) <- NULL
```{r}
B
rownames(B) <- NULL
colnames(B) <- NULL
```

How to replace specifc element of a matrix:
let's say you wish to change a value
step1: check existing value
```{r}
B[2,2]
```
step2: change value
```{r}
B[2,2] <- 211
```
step3: check if matrix has changed
```{r}
B
```














